\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-04 fully-abstract-comp/popl2019/paper.tex
Formatting of figure
Added outline for section about target machine.
Some work on target machine section.

\input{../preamble}
% Paper specific redefinitions of commands
\renewcommand{\MemSeg}{\shareddom{MemSeg}}
\renewcommand{\Reg}{\shareddom{RegFile}}
\renewcommand{\RegName}{\shareddom{RegName}}
\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
\newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}

@@

@@

\section{A Capability Machine with Sealing and Linear Capabilities}
\label{sec:cap-mach-w-seal-and-lin}
\begin{itemize}
\item present our \emph{target language} and its operational semantics (excerpts)
\item mention roughly what components look like
\end{itemize}
\begin{figure}[htb]
  \centering
  \[
  \arraycolsep=1.4pt
  \begin{array}{rrcl l rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
    &\aendbnf & \in & \Addr \cup \{\infty \} & & &\sealendbnf & \in & \Seal \cup \{\infty \}\\
    \scbnf \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
    \cbnf \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
    \rbnf \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
    &\Reg & \defeq & \RegName \fun \Word & & &\Mem & \defeq & \Addr \fun \Word\\
    &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
    &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
  \end{array}
\]
\[
  \arraycolsep=1.4pt
\begin{array}{rcl}
\multicolumn{3}{l}{    \arraycolsep=0pt
      \com{r} \in  \tRegName \hspace{2.5cm}   \com{\rn} \defbnf \com{r} \mid \nats
}\\
\Instr &\defbnf & \tfail \mid \thalt \mid \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}}\mid\tgetp{\com{r}}{\com{r}} \mid\\
      & &    \tgetlin{\com{r}}{\com{r}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tstore{\com{r}}{\com{r}} \mid \tload{\com{r}}{\com{r}} \mid \tcca{\com{r}}{\com{\rn}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \\
  & & \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tsetatob{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid \tcseal{\com{r}}{\com{r}} \mid \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\ 
      & & \tsplice{\com{r}}{\com{r}}{\com{r}} 
\end{array}
\]
  \caption{The syntax of our capbility machine with seals and linera capbilities.}
  \label{fig:target-syntax}
\end{figure}

@@

@@

@@

>>>>

\input{../preamble}
% Paper specific redefinitions of commands
\renewcommand{\MemSeg}{\shareddom{MemSeg}}
\renewcommand{\Reg}{\shareddom{RegFile}}
\renewcommand{\RegName}{\shareddom{RegName}}
\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
\renewcommand{\perm}{\var{p}}
\renewcommand{\SealableCaps}{\shareddom{Sealables}}
\renewcommand{\Cap}{\shareddom{NatTok}}

\newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}

@@

\newcommand{\trgcm}{\textsc{LCM}}
\newcommand{\srccm}{\textsc{SLCM}}

@@

\section{A Capability Machine with Sealing and Linear Capabilities}
\label{sec:cap-mach-w-seal-and-lin}
\begin{figure}[tb]
  \centering
  \[
  \arraycolsep=1.4pt
  \begin{array}{rrcl l rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
    \aendbnf \in & \Addr \cup \{\infty \} & & & & \sealendbnf \in & \Seal \cup \{\infty \}& &\\
    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
    \vsc \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
    c \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
    r \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
    &\Reg & \defeq & \RegName \fun \Word & & &\Mem & \defeq & \Addr \fun \Word\\
    &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
    &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
  \end{array}
\]
\[
  \arraycolsep=1.4pt
\begin{array}{rcl}
\multicolumn{3}{l}{    \arraycolsep=0pt
      \com{r} \in  \tRegName \hspace{2.5cm}   \com{\rn} \defbnf \com{r} \mid \nats
}\\
\Instr &\defbnf & \tfail \mid \thalt \mid \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}}\mid\tgetp{\com{r}}{\com{r}} \mid\\
      & &    \tgetlin{\com{r}}{\com{r}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tstore{\com{r}}{\com{r}} \mid \tload{\com{r}}{\com{r}} \mid \tcca{\com{r}}{\com{\rn}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \\
  & & \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tsetatob{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid \tcseal{\com{r}}{\com{r}} \mid \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\ 
      & & \tsplice{\com{r}}{\com{r}}{\com{r}} 
\end{array}
\]
  \caption{The syntax of our capbility machine with seals and linera capbilities.}
  \label{fig:target-syntax}
\end{figure}
% Brief intro to section and this cap machine:
% Cap machine inspired by CHERI, but with linear capabilities
In this section, we introduce a simple but representative capability machine with linear capabilities.
We will later introduce a capability machine similar to the one in this section, so to easily distinguish the two, we will call this one \trgcm{} (Linear Capability Machine).
\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main exception.
In order to avoid details uninteresting to the focus of this work, \trgcm{} assumes an infinite address and unbounded integers.
\lau{Do we have a good reason to assume an infinite address space for this work?
  We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}

\begin{wrapfigure}{r}{0.33\linewidth}
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (rwx) {$\rwx$};
    \node[main node] (rx) [below right of=rwx] {$\rx$};

    \node[main node] (rw) [below left of=rwx,] {$\rw$};
    \node[main node] (r) [below right of=rw] {$\readonly$};
    \node[main node] (0) [below of=r] {$\noperm$};

    \path[every node/.style={font=\sffamily\small}]

    (rw) edge (r)
    (r) edge (0)

    (rwx) edge (rx)

    (rw) edge (rwx)
    (r) edge (rx);
  \end{tikzpicture}

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{wrapfigure}
%%% Capabilities
The concept of a capability is the cornerstone of any capability machine.
In its simplest form, a capability is a permission and a range of authority.
The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
% Permissions and Range of authority
The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}) where $\var{perm}$ is the permission and $[\var{base},\var{end}]$ is the range of authority.
\trgcm{} only has memory capabilities, and we limit them to a number of sensible permissions making the available permissions: read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$).
We define the preorder $\le$ on permissions as illustrated in Figure~\ref{fig:perm-hier}.
% Pointer and linearity
In addition to the two basic things, the capabilities also have a current address $\addr$ and a linearity $\lin$. The linearity of a capability can either be $\linear$ or $\normal$ where $\normal$ corresponds to a traditional capability. We will describe linear capabilities in detail when we get to the operational semantics of \trgcm{}.

%%% Seals
Any reasonable capability machine needs a way to setup boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
On \trgcm{} we have sealed capabilities to achieve this.
Syntactically, sealed capabilities add two things to the capability machine: Sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ and sealed capabilities $\sealed{\sigma,\vsc}$.
% seals
A set of seals consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$ that is represented by a natural number.
% sealed caps
A sealed capability consists of a sealable $\vsc$ and the seal $\sigma$ it has been sealed with.
The sealables are sets of seals and capabilities, and we refer to the set of them as $\SealableCaps$.
The system native tokens on \trgcm{} are the sealables and the sealed capabilities and we refer to this set as $\Cap$.

%%% Domains
The words on \trgcm{} are the system native tokens in $\Cap$ and data which is represented by $\ints$.
We assume a finite set $\RegName$ of register names that at least contains $\{\pcreg , \rretd , \rretc , \rstk , \rdata , \rtmp{1} , \rtmp{2}\}$, and we define register files as a function from register names to words.
A memory is a full memory and a memory segment is part of a memory.
Both are defined as functions from addresses, defined as the natural numbers, to words, but the memory segment functions only need to be partial.
$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution with an attempt to break the capability machine guarantees.
An executable configuration is a register file and memory pair.

% syntax (instructions)
% opsem
% move w/ lin
% load w/ lin
% cseal
% xjmp
% split/splice

% Components ...
% write xor exec
% contents
% linking

@@

    \begin{array}{l}

@@

  \end{cases}\\

@@

  \end{array}
